---
title: "Map Spisula Sites"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Intro and Practice

The code used here for reference is from Harmony, Honggang (https://github.com/hzz0024/BioinfoTools), and https://geocompr.robinlovelace.net/adv-map.html.

My goal is to make a map of the spisula sample sites from 2012 to 2020, with number of samples coded by the size of the circle and subspecies coded by color.

```{r lib, echo=FALSE, message=FALSE}
library(sf)
library(raster)
library(dplyr)
library(spData)
#library(spDataLarge)

library(tmap)
#library(leaflet) # for interactive maps
library(ggplot2)
```

``` {r nztest, eval=FALSE}
tm_shape(nz) +
    tm_fill() 
map_nz1 = tm_shape(nz) + tm_polygons()
#add elevation
#map_nz1 = map_nz +
#  tm_shape(nz_elev) + tm_raster(alpha = 0.7) #need to define nz_elevation
#representing nz territorial waters
nz_water = st_union(nz) %>% st_buffer(22200) %>% 
  st_cast(to = "LINESTRING")
map_nz2 = map_nz1 +
  tm_shape(nz_water) + tm_lines()
#add peaks
map_nz3 = map_nz2 +
  tm_shape(nz_height) + tm_dots()  #but why don't I need to define nz_height
tmap_arrange(map_nz1, map_nz2, map_nz3)
```

```{r regional, echo=FALSE, message=FALSE}
#Experimenting with tmap
#All installs are removed to knit properly
library(sf)
library(raster)
library(dplyr)
library(spData)
#library(spDataLarge) # Not for current version of R

library(tmap)
#library(leaflet) # for interactive maps
library(ggplot2)

tm_shape(us_states) + tm_graticules() + tm_polygons()

#create a region in nz
nz_region = st_bbox(c(xmin = 1340000, xmax = 1450000,
                      ymin = 5130000, ymax = 5210000),
                    crs = st_crs(nz_height)) %>% st_as_sfc()

nz_height_map = tm_shape(nz_region, bbox = nz_region) +
  tm_raster(style = "cont", palette = "YlGn", legend.show = TRUE) +
  tm_shape(nz_height) + tm_symbols(shape = 2, col = "red", size = 1) +
  tm_scale_bar(position = c("left", "bottom"))

tm_shape(nz_region) + tm_polygons()
tm_shape(nz) + tm_polygons() + tm_grid()

#try with usa
tm_shape(us_states) + tm_polygons() + tm_grid()
eastcoast=st_bbox(c(xmin = -85, xmax = -65,
                      ymin = 30, ymax = 45), 
                  crs = st_crs(us_states)) %>% st_as_sfc()
#WOOOHOOOO
#Shape = us states, box = where I defined it above
tm_shape(us_states, bbox = eastcoast) + tm_graticules() + tm_polygons()

#try with world instead of us so that canada doesn't look like water
#tm_shape(world, bbox = eastcoast) +  
#  tm_fill(col = "seashell1") + tm_borders(col="grey10") + tm_layout(bg.color = "paleturquoise1") +
#  tm_graticules()
# Works but then there are no state lines, and it is bulkier
```

```{r formatit}
library(sf)
library(raster)
library(dplyr)
library(spData)
library(tmap)
library(ggplot2)
library(tidyverse)

#Show full map
tm_shape(us_states) + tm_polygons() + tm_graticules() 
#create my bounds with st_bbox, piped into st as sfc
eastcoast=st_bbox(c(xmin = -85, xmax = -65,
                      ymin = 30, ymax = 45)) %>% st_as_sfc()
#use those bounds under bbox variable within tm_Shape
tm_shape(us_states, bbox = eastcoast) + tm_polygons() + tm_graticules() 
#tm_polygons = tm_fill + tm_borders
#color and formatting
tm_shape(us_states, bbox = eastcoast) +  
  tm_fill(col = "grey99") + tm_borders(col="grey30") + tm_layout(bg.color = "grey90") +
  tm_graticules(col="white")
```
## Importting Data

```{r importdata, message=FALSE}

spisloc <- read_csv("/Users/hannah/gitHub/Spisula/Work/map_spisula_input.csv")

spisloc %>%
  ggplot() +
  geom_point(aes(x=Lat, y=Lon))

#Add on to map
#example from: https://www.r-bloggers.com/2019/10/make-a-map-of-your-study-site-with-tmap/
sitescoral <- read.csv(url("https://raw.githubusercontent.com/cbrown5/BenthicLatent/master/data-raw/JuvUVCSites_with_ReefTypes_16Jun2016.csv"))
sites <- st_as_sf(sitescoral, coords = c("coordx", "coordy"))
#making the coordinate reference system the same
#land <- st_read(dsn ="LandPoly")
#st_crs(sites) <- st_crs(land)

  
#my data
spissites <- st_as_sf(spisloc, coords = c("Lon","Lat")) #LON FIRST - if confused, view whole world
#making the coordinate reference system the same
st_crs(spissites) <- st_crs(us_states)


library(RColorBrewer)
#plotting
tm_shape(us_states, bbox=eastcoast) +  
  tm_fill(col = "grey99") + tm_borders(col="grey30") + tm_layout(bg.color = "grey90") +
  tm_grid(col="white") +
  tm_shape(spissites) +  
  tm_symbols(size = 0.9, col = "Species", alpha = 0.5,  palette = "Set2") 
#For the future: to reverse a pallette, easy as: palette = rev("Set2")
```

### Final Adjustments

Possible iterations of the graph:
+Where color or shape indicate species
+Whiter background
+Where size is number of samples taken at the site
+With nantuket similis offset

```{r finaladjust}
eastcoast=st_bbox(c(xmin = -85, xmax = -65,
                      ymin = 30, ymax = 45)) %>% st_as_sfc()
#Base
map_1 <- tm_shape(us_states, bbox=eastcoast) +  
  tm_fill(col = "grey99") + tm_borders(col="grey30") + tm_layout(bg.color = "grey90") +
  tm_graticules(col="white") +
  tm_shape(spissites) +  
  tm_symbols(size = 1, col = "Species", alpha = 0.5,  palette = "Set2", border.alpha = 0.2)
#Size = number of samples
map_2 <- tm_shape(us_states, bbox=eastcoast) +  
  tm_fill(col = "grey99") + tm_borders(col="grey30") + tm_layout(bg.color = "grey90") +
  tm_graticules(col="white") +
  tm_shape(spissites) +  
  tm_symbols(size = "Samples", col = "Species", alpha = 0.5,  palette = "Set2", border.alpha = NA)
#Also changed border alpha to default
#All are too small!!

#By shapes
map_3 <- tm_shape(us_states, bbox=eastcoast) +  
  tm_fill(col = "grey99") + tm_borders(col="grey30") + tm_layout(bg.color = "grey90") +
  tm_graticules(col="white") +
  tm_shape(spissites) +  
  tm_symbols(size = 1, shape = "Species",  alpha = 0.7,  col= "Samples", palette = "Reds")

#Try to change size and jitter (for sites with both)
map_4 <- tm_shape(us_states, bbox=eastcoast) +  
  tm_fill(col = "grey99") + tm_borders(col="grey30") + tm_layout(bg.color = "grey90") +
  tm_graticules(col="white") +
  tm_shape(spissites) +  
  tm_symbols(size = "Samples", col = "Species", alpha = 0.7,  palette = "Set2",
             scale = 2.5, jitter = 0.01) 
#default icon.scale=3 ,  icon.scale = 9 has no change
#default legend.max.symbol.size = 0.8 , on its own legend.max.symbol.size = 3 has no effect
#THIS ONE: default scale=1, changing this makes them all bigger

#Circles not filled in
map_5 <- tm_shape(us_states, bbox=eastcoast) +  
  tm_fill(col = "grey99") + tm_borders(col="grey30") + tm_layout(bg.color = "grey90") +
  tm_graticules(col="white") +
  tm_shape(spissites) +  
  tm_symbols(size = "Samples", alpha = 1, col = "Species",shape=1, border.lwd = 2, palette = "Set2",
             scale = 2.5, jitter = 0.01) 
#Do so by choosing a different shape (hollow circle) rather than trying to set border color to variable

#map_1 at larger scale
map_6 <- tm_shape(us_states, bbox=eastcoast) +  
  tm_fill(col = "grey99") + tm_borders(col="grey30") + tm_layout(bg.color = "grey90") +
  tm_graticules(col="white") +
  tm_shape(spissites) +  
  tm_symbols(size = 1, jitter= 0.01 ,col = "Species",  palette = "Set2",alpha = 0.6,border.lwd = 0.7,  border.alpha = 0.5, scale = 1.5)

map_7 <- tm_shape(us_states, bbox=eastcoast) +  
  tm_fill(col = "grey99") + tm_borders(col="grey30") + tm_layout(bg.color = "grey94") +
  tm_graticules(col="white") +
  tm_shape(spissites) +  
  tm_symbols(size = 1, jitter= 0.1, shape=1 ,col = "Species",  palette = "Set2",border.lwd = 2, scale = 1.5)

#tmap_arrange(map_1,map_2,map_3)

#Base
map_1
#By size too small
map_2
#By shape with quantity as color
map_3
#Scaled up by size and miniscule jitter
map_4
#Circles scaled and jitter
map_5
#Base redone with scaling and jitter
map_6
#Base with empty circles
map_7
#Filled in small circles
map_1 <- 2

tm_shape(us_states, bbox=eastcoast) +  
  tm_fill(col = "grey99") + tm_borders(col="grey30") + tm_layout(bg.color = "grey90") +
  tm_grid(col="white") +
  tm_shape(spissites) +  
  tm_symbols(size = 0.2, jitter=0.05,col = "Species",  palette = "Set2") 

```


```{r tryggmap, include=FALSE, message=FALSE, eval=FALSE}
library(lubridate)
library(ggplot2)
library(dplyr)
library(data.table)
library(ggrepel)
library(tidyverse)

#Example from https://www.littlemissdata.com/blog/maps

incidents= read.csv('https://raw.githubusercontent.com/lgellis/MiscTutorial/master/ggmap/i2Sample.csv', stringsAsFactors = FALSE)
n <- read.csv('https://raw.githubusercontent.com/lgellis/MiscTutorial/master/ggmap/n.csv', stringsAsFactors = FALSE)

col1 = "#011f4b"
col2 = "#6497b1"
col3 = "#b3cde0"
col4 = "#CC0000"

incidents$ymd <- mdy_hms(incidents$Event.Clearance.Date)
incidents$year <- year(incidents$ymd)

i2 <- incidents %>% 
  filter(year>=2017 & year<=2018)
i2[complete.cases(i2), ]

n$label <-paste(n$Rank, n$Location, sep="-")

#if(!requireNamespace("devtools")) install.packages("devtools")
#devtools::install_github("dkahle/ggmap", ref = "tidyup", force=TRUE)
library("ggmap")
ggmap::register_google(key = "AIzaSyAVHjpzedI9cV034dq71k1-BZIRonPwEBo")

#Code from: https://cran.r-project.org/web/packages/ggmap/readme/README.html
eastcoast <- c(left = -85, bottom = 30, right = -65, top = 45)
#zoom is how accurate it is. Zoom 10 looks great but takes forever, Zoom 5 is too sloppy
#also impacts the number of labels, more zoomed in has no state labels, but all the other labels are real tiny.
coastmap <- get_stamenmap(eastcoast, zoom = 6, maptype = "toner-lite") %>% ggmap() 

coastmap

#ggmap(get_stamenmap(bbox = c(left = -95.80204, bottom = 29.38048, right =
  -94.92313, top = 30.14344), zoom = 10, maptype = c("terrain",
  "terrain-background", "terrain-labels", "terrain-lines", "toner",
  "toner-2010", "toner-2011", "toner-background", "toner-hybrid",
  "toner-labels", "toner-lines", "toner-lite", "watercolor"),
  crop = TRUE, messaging = FALSE, urlonly = FALSE,
  color = c("color", "bw"), force = FALSE))


#get_stamenmap(eastcoast, zoom = 6, maptype = "terrain")            %>% ggmap() # google maps basic
#get_stamenmap(eastcoast, zoom = 6, maptype = "terrain-background", color ="bw", force=TRUE) %>% ggmap() # terrain with no labels
# get_stamenmap(eastcoast, zoom = 6, maptype = "terrain-labels")     %>% ggmap() # just labels
# get_stamenmap(eastcoast, zoom = 6, maptype = "terrain-lines")      %>% ggmap() # dotted lines except coastline
# get_stamenmap(eastcoast, zoom = 6, maptype = "toner")              %>% ggmap() # labels, ocean is black
# get_stamenmap(eastcoast, zoom = 6, maptype = "toner-2010")         %>% ggmap() # same
# get_stamenmap(eastcoast, zoom = 6, maptype = "toner-2011")         %>% ggmap() # same
#get_stamenmap(eastcoast, zoom = 6, maptype = "toner-background", color="color", force=TRUE)   %>% ggmap() # works but change color
# get_stamenmap(eastcoast, zoom = 6, maptype = "toner-hybrid")       %>% ggmap() # no coast lines, mess
# get_stamenmap(eastcoast, zoom = 6, maptype = "toner-labels")       %>% ggmap() # just labels
# get_stamenmap(eastcoast, zoom = 6, maptype = "toner-lines")        %>% ggmap() # just lines except costline
get_stamenmap(eastcoast, zoom = 9, maptype = "toner-lite")         %>% ggmap() # includes placenames
# get_stamenmap(eastcoast, zoom = 6, maptype = "watercolor")         %>% ggmap() # silly colors

#THE THING I WANT DOES NOT EXIST: https://github.com/stamen/toner-carto/issues/16

```

